---
title: "p8105_hw6_cx2347"
author: "Chuyuan XU"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(stringr)
library(broom)
library(patchwork)
library(rvest)
library(httr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.9,
  out.width = "90%"
)
```

### Problem 1
```{r p1 import, message=FALSE}
csv = "https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv"

homicide_raw = read_csv(csv, na=c("NA", "", "."))
```

```{r p1 tidy}
homicide_df = homicide_raw |>
  janitor::clean_names() |>
  
  # Create a city_state variable (e.g. “Baltimore, MD”)
  mutate(city_state = str_c(city, ', ', state)) |>
  
  # omit Dallas, TX; Phoenix, AZ; Kansas City, MO; Tulsa, AL
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
  ) |>
  
  # limit the data among those for whom victim_race is white or black
  filter(
    victim_race %in% c("White", "Black")
  ) |>
  
  # change the victims' age into numbers
  # change the unknown victims' sex into NA
  # change the victim_race into a factor
  # create a indicator for resolved vs unresolved
  mutate(
    victim_age = if_else(victim_age == "Unknown", NA , victim_age),
    victim_age = as.numeric(victim_age),
    victim_sex =  if_else(victim_sex == "Unknown", NA , victim_sex),
    victim_race = fct_relevel(victim_race, "White"),
    resolved = as.numeric(disposition == "Closed by arrest")
  )
```

```{r, p1 glm_Balt}
# fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors
fit_balt = homicide_df |>
  filter(city == "Baltimore") |>
  glm(
    resolved ~ victim_age + victim_race + victim_sex, 
    data = _,
    family = binomial()
  )

fit_balt_df = fit_balt |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate),
    conf.low = exp(estimate - 1.96*std.error),
    conf.high = exp(estimate + 1.96*std.error),
  ) |>
  select(term, OR, conf.low, conf.high, p.value) 

fit_balt_df |> 
  mutate(
    conf.low = round(conf.low, 3),
    conf.high = round(conf.high, 3)
  ) |>
  nest(CI = c("conf.low", "conf.high")) |>
  select(term, OR, CI, p.value) |>
  knitr::kable(
    digits = 3,
    col.names = c("terms", "est OR", "95% CI (OR)", "P-value"),
    align = 'r'
  )
```

The estimated adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is `r round(pull(filter(fit_balt_df, term == "victim_sexMale"), OR),3)`, with a 95% confidence interval of (`r round(pull(filter(fit_balt_df, term == "victim_sexMale"), conf.low), 3)`, `r round(pull(filter(fit_balt_df, term == "victim_sexMale"), conf.high),3)`).

```{r p1 glm_all cities}
glm_cities_df = homicide_df |>
  nest(data = -city_state) |>
  mutate(
    models = map(data, \(df) glm(resolved ~ victim_age + victim_race + victim_sex, data = df)),
    results = map(models, broom::tidy)
  ) |>
  select(-data, -models) |> 
  unnest(results) |>
  
  # keep the results for only adj OR males over females
  filter(term == "victim_sexMale") |>
  
  mutate(
    OR = exp(estimate),
    conf.low = exp(estimate - 1.96*std.error),
    conf.high = exp(estimate + 1.96*std.error)
  ) |> 
  select(city_state, OR, conf.low, conf.high, p.value) |>
  
  # Organize cities according to estimated OR
  mutate(
    city_state = fct_reorder(city_state, OR)
  ) |>
  arrange(city_state)
```

```{r p1 plot}
# Create a plot that shows the estimated ORs and CIs for each city.
glm_cities_df |>
  mutate(
    fill = (OR > 1)
  ) |>
  ggplot(aes(y = city_state, x = OR, color = fill)) +
  geom_point() +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    width = 0.2,
    alpha = 0.8
  ) +
  geom_vline(
    aes(xintercept = 1.0), 
    color = "grey",
    alpha = 0.8
  ) +
  theme_minimal() +
  labs(
    y = 'City, State',
    x = 'the Estimated OR with 95% CI',
    color = 'OR is greater than 1.0',
    title = 'the Adjusted Odds Ratio (and 95% CI) for Solving Homicides',
    subtitle = 'Comparing Male Victims to Female Victims'
  )
```

#### Comments:
6 of 47 cities in the study have an OR greater than 1, for solving homicides comparing male and female victims. In most cities, the odds for solving homicides with a male victims is lower than that for solving homicides with a female victims, adjusting for their age and race (white or black). However, many 95% CI of the adjusted OR include the null value of 1, we cannot conclude those associations are significant.


### Problem 2
```{r p2 import}
data("weather_df")
```

```{r p3 bootstrap}
set.seed(1234)

weather_bs = weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df))
  )

# get the r_sqr
weather_rsq = weather_bs |>
  mutate(
    results = map(models, broom::glance)
  ) |>
  unnest(results) |>
  mutate(
    id = as.numeric(.id)
  ) |>
  select(id = id, r.squared)

# get the beta_1/beta_2
weather_betas = weather_bs |>
  mutate(
    results = map(models, broom::tidy)
  ) |>
  select(.id, results) |>
  unnest(results) |>
  select(.id, term, estimate) |>
  pivot_wider(
    names_from = "term",
    values_from = "estimate",
    id_cols = .id
  ) |>
  mutate(
    beta.frac = tmin/prcp,
    id = as.numeric(.id)
  ) |>
  select(id, beta.frac)

# merge the result into one
weather_bs_rslt = full_join(weather_rsq, weather_betas, by = "id")
```

```{r p2 dist plot}
rsqr.dist = weather_bs_rslt |>
  ggplot(aes(x = r.squared)) +
  geom_density(fill = 'grey', alpha = 0.4) +
  geom_vline(aes(xintercept = mean(r.squared)), color = 'red', linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = median(r.squared)), color = 'blue', linetype = "solid", alpha = 0.8) +
  labs(
    x = "R sqaured"
  ) +
  theme_minimal()

beta.frac_dist = weather_bs_rslt |>
  ggplot(aes(x = beta.frac)) +
  geom_density(fill = 'grey', alpha = 0.4) +
  geom_vline(aes(xintercept = mean(beta.frac)), color = 'red', linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = median(beta.frac)), color = 'blue', linetype = "solid", alpha = 0.8) +
  labs(
    x = "Beta_tmin/Beta_prcp",
    y = ""
  ) +
  theme_minimal()

rsqr.dist + beta.frac_dist +
  plot_annotation(
    title = 'the Distribution of the Estimates',
    caption = 'the red dashed line is the mean and the blue solid line is the median')
```

The above figure displays the distribution of Estimated Beta_tmin/Beta_prcp and R-squared, where the red dashed line is the mean and the blue solid line is the median. The r-squared is approximately symmetrically distributed with a range of 0.930 and 0.955. The Beta_tmin/Beta_prcp r-squared has a left-skewed distribution, with a range between -450 and -100.

```{r p2 quantiles}
q = weather_bs_rslt |>
  pivot_longer(
    cols = r.squared:beta.frac,
    values_to = "estimate",
    names_to = "terms"
  ) |>
  group_by(terms) |>
  summarize(
    ci_lower = quantile(estimate, 0.025), 
    ci_upper = quantile(estimate, 0.975)
  ) |>
  mutate(
    terms = case_match(
      terms,
      "beta.frac" ~ "est(Beta_tmin/Beta_prcp)",
      "r.squared" ~ "R-squared"
      )
    )

q |>
  knitr::kable(
    digits = 3,
    col.names = c("Estimates", "2.5% quantiles", "97.5% quantiles"),
    caption = "the 95% Confidence Intervals for Estimates"
  )
```

The above table shows the 95% Confidence Intervals for the Estimated Beta_tmin/Beta_prcp and R-squared.

### Problem 3
```{r p3 import}
wt_raw = read_csv("data/birthweight.csv", na = c("NA", ".", ""))
```

```{r p3 tidy}
wt_df = wt_raw |>
  janitor::clean_names() |>
  mutate(
    babysex = case_match(
      babysex,
      1 ~ "male",
      2 ~ "female"
    ),
    malform = case_match(
      0 = "absent", 
      1 = "present"
    ),
    
    frace = case_match(
      frace,
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other", 
      9 ~ NA
    ),
    
    mrace = as.factor(
      frace,
      1 ~ "White", 
      2 ~ "Black", 
      3 ~ "Asian", 
      4 ~ "Puerto Rican", 
      8 ~ "Other", 
      9 ~ NA
    )
  )

```


