---
title: "p8105_hw6_cx2347"
author: "Chuyuan XU"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(stringr)
library(broom)
library(patchwork)
library(rvest)
library(httr)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.9,
  out.width = "90%"
)
```

### Problem 1
```{r p1 import, message=FALSE}
csv = "https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv"

homicide_raw = read_csv(csv, na=c("NA", "", "."))
```

```{r p1 tidy}
homicide_df = homicide_raw |>
  janitor::clean_names() |>
  
  # Create a city_state variable (e.g. “Baltimore, MD”)
  mutate(city_state = str_c(city, ', ', state)) |>
  
  # omit Dallas, TX; Phoenix, AZ; Kansas City, MO; Tulsa, AL
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
  ) |>
  
  # limit the data among those for whom victim_race is white or black
  filter(
    victim_race %in% c("White", "Black")
  ) |>
  
  # change the victims' age into numbers
  # change the unknown victims' sex into NA
  # change the victim_race into a factor
  # create a indicator for resolved vs unresolved
  mutate(
    victim_age = if_else(victim_age == "Unknown", NA , victim_age),
    victim_age = as.numeric(victim_age),
    victim_sex =  if_else(victim_sex == "Unknown", NA , victim_sex),
    victim_race = fct_relevel(victim_race, "White"),
    resolved = as.numeric(disposition == "Closed by arrest")
  )
```

```{r, p1 glm_Balt}
# fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors
fit_balt = homicide_df |>
  filter(city == "Baltimore") |>
  glm(
    resolved ~ victim_age + victim_race + victim_sex, 
    data = _,
    family = binomial()
  )

fit_balt_df = fit_balt |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate),
    conf.low = exp(estimate - 1.96*std.error),
    conf.high = exp(estimate + 1.96*std.error),
  ) |>
  select(term, OR, conf.low, conf.high, p.value) 

fit_balt_df |> 
  mutate(
    conf.low = round(conf.low, 3),
    conf.high = round(conf.high, 3)
  ) |>
  nest(CI = c("conf.low", "conf.high")) |>
  select(term, OR, CI, p.value) |>
  knitr::kable(
    digits = 3,
    col.names = c("terms", "est OR", "95% CI (OR)", "P-value"),
    align = 'r'
  )
```

The estimated adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is `r round(pull(filter(fit_balt_df, term == "victim_sexMale"), OR),3)`, with a 95% confidence interval of (`r round(pull(filter(fit_balt_df, term == "victim_sexMale"), conf.low), 3)`, `r round(pull(filter(fit_balt_df, term == "victim_sexMale"), conf.high),3)`).

```{r p1 glm_all cities}
glm_cities_df = homicide_df |>
  nest(data = -city_state) |>
  mutate(
    models = map(data, \(df) glm(resolved ~ victim_age + victim_race + victim_sex, data = df)),
    results = map(models, broom::tidy)
  ) |>
  select(-data, -models) |> 
  unnest(results) |>
  
  # keep the results for only adj OR males over females
  filter(term == "victim_sexMale") |>
  
  mutate(
    OR = exp(estimate),
    conf.low = exp(estimate - 1.96*std.error),
    conf.high = exp(estimate + 1.96*std.error)
  ) |> 
  select(city_state, OR, conf.low, conf.high, p.value) |>
  
  # Organize cities according to estimated OR
  mutate(
    city_state = fct_reorder(city_state, OR)
  ) |>
  arrange(city_state)
```

```{r p1 plot}
# Create a plot that shows the estimated ORs and CIs for each city.
glm_cities_df |>
  mutate(
    fill = (OR > 1)
  ) |>
  ggplot(aes(y = city_state, x = OR, color = fill)) +
  geom_point() +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    width = 0.2,
    alpha = 0.8
  ) +
  geom_vline(
    xintercept = 1.0,
    color = "grey",
    alpha = 0.8
  ) +
  theme_minimal() +
  labs(
    y = 'City, State',
    x = 'the Estimated OR with 95% CI',
    color = 'OR is greater than 1.0',
    title = 'the Adjusted Odds Ratio (and 95% CI) for Solving Homicides',
    subtitle = 'Comparing Male Victims to Female Victims'
  )
  
```

#### Comments:
6 of 47 cities in the study have an OR greater than 1, for solving homicides comparing male and female victims. In most cities, the odds for solving homicides with a male victims is lower than that for solving homicides with a female victims, adjusting for their age and race (white or black). However, many 95% CI of the adjusted OR include the null value of 1, we cannot conclude those associations are significant.


### Problem 2
```{r}

```

